Active Object Search with a Mobile Device\\for People with Visual Impairments

### 2018-12-28 review_edition/reviewed.tex
final revision

\keywords{Active vision, object search, vision impairment, Markov Decision Process}

\abstract{Modern smartphones can provide a multitude of services to assist people with visual impairments, and their cameras in particular can be useful for assisting with tasks, such as reading signs or searching for objects in unknown environments. Previous research has looked at ways to solve these problems by processing the camera's video feed, but very little work has been done in {\em actively} guiding the user towards specific points of interest, maximising the effectiveness of the underlying visual algorithms. In this paper, we propose a control algorithm based on a Markov Decision Process that uses a smartphone's camera to generate real-time instructions to guide a user towards a target object. The solution is part of a more general active vision application for people with visual impairments. An initial implementation of the system on a smartphone was experimentally evaluated with participants with healthy eyesight to determine the performance of the control algorithm. The results show the effectiveness of our solution and its application to help people with visual impairments find objects in unknown environments.}

@@

\noindent It is estimated that almost half a billion people worldwide live with mild to severe visual impairments or total blindness~\cite{bourne2017magnitude} and significant effort is being made to enable these people to lead more independent lives. Modern improvements in mobile computing power and image processing techniques have provided researchers with new and powerful tools to solve this problem. The work presented here is part of a project to assist people with vision impairments to navigate and find objects in unknown environments with the aid of a smartphone. The proposed system implements ideas from the field of active vision~\cite{Bajcsy2017}, but replaces the typical electro-mechanical actuators of a moving camera with the body (i.e.\ arm, hand) of the user holding the smartphone, as pictured in Figure~\ref{fig:system-use}, expanding upon concepts originally proposed in~\cite{bellotto2013} and~\cite{lock2017portable}.

@@

\noindent Assistive technology for people living with vision impairments is a growing research area~\cite{Manduchi2012,khoo2016multimodal}. In recent years, the increase in mobile processing power and computer vision improvements have led to research in the use of smartphone cameras to augment or enhance a user's vision and help them find objects or other points of interest. Earlier attempts at the problem involved placing special markers or barcodes around an environment, which the user then scans with a smartphone or similar mobile device~\cite{gude2013blind,iannizzotto2005badge3d,manduchi2012mobile}. This device then uses some feedback mode, e.g.\ Braille or sound, to guide the user towards the target. %The major drawbacks to this approach is the significant upfront effort required to augment entire areas with markers to a point where the system will be useful to a VI user. Furthermore, these passive systems rely on the user scanning the marker in the first place; something that can be made more complicated with markers degrading over time, changing ambient lighting, different viewing angles and marker occlusion.

Another approach is to discard tags completely and rely on computer vision to perform the object detection, something that has become more practical with recent improvements to feature detectors and deep networks~\cite{huang2017speed,redmon2016you}. SIFT and SURF-based object detectors have also been used to detect known objects, when they are in the camera's view, and to guide the user to them using sonified instructions~\cite{schauerte2012assistive}. These type of systems is more flexible than the tag-based ones, but it has the same drawback of being {\em passive}, in the sense that it relies on having the object within the camera's view in the first place. Also, no clear performance metrics are reported in the previous paper. The VizWiz system~\cite{bigham2010vizwiz} offloads the object recognition tasks to an Amazon Mechanical Turk worker who then provides feedback on where the object of interest is located relative to the user. The VizWiz has the advantage of being fairly robust and is able to classify a great deal of objects with little effort from the user and can provide natural, human-generated and curated directions. However, this approach does not enhance user independence, since a person with vision impairments is now beholden to an online worker instead of a relative, friend or bystander. Furthermore, a good internet connection is required on the device, possibly limiting its use in some poor-reception areas.

@@

In summary, much research has been conducted on recognition of and guidance towards target objects, including active vision solutions for image classifiers and robotic systems. However, to our knowledge, no previous work has been done on {\em active} object search and guidance for humans, which would especially benefit people with vision impairments. In this paper, we implement such an active vision system with a human in the loop
% ~\cite{bellotto2013,lock2017portable}
that guides the user towards an out-of-view target object. Our system exploits prior knowledge of the objects spatial distribution within an indoor environment, learned from a dataset of real-world images, and the history of past object observations made during the search.
%With this, we hope to make the searching process less random and move towards a more natural and efficient object searching process to the benefit of the VI\@.

@@

\noindent The work presented in this paper is a fundamental step towards a more general project's goal to develop a stand-alone system that can guide a person with vision impairments to his/her destination with minimal user input or intervention. A complete system diagram is given is Figure~\ref{fig:control-loop}. This closed-loop system is conceptually similar to other classical control problems, where the difference between desired and actual state of a process is used to generate a control signal that changes the process itself.

@@

In a real application for people with visual impairments, the position of the waypoint would be given to the user by a set of audio or vibrotactile instructions. However, since we are mainly interested in evaluating the control algorithm of our system and not the interface ($K$ and not $u$), our current prototype generates guidance instructions with four on-screen arrows (see Figure~\ref{fig:system-screenshot}).
% This might seem like an unnecessary extra step, but this system is designed for user with vision impairments who are limited to audio instructions, which makes it necessary to split the instructions up into separate channels that can be communicated more simply.
% This requirement becomes clear when one considers the bandwidth limitations of a person's audio channel compared to the visual one.
Obviously, this visual interface is only used for debugging and experimental evaluation of the controller, and it will be replaced by an opportune audio interface, e.g.~\cite{bellotto2013}, at a second stage.

@@

Table~\ref{tab:results-full} shows a fairly consistent spread for the TAR across the participants. The inter-participant spread ($\sigma=11\%$) is fairly significant, perhaps indicating that the user's search behaviour and strategy affects the target acquisition performance, but with an average TAR of $82\%$, it is clear that the system successfully finds the target object during the vast majority of searches.

@@

\noindent The cumulative distribution of the time it took the participants to reach the target object is given in Figure~\ref{fig:time-participants}. We see that the distribution is heavily skewed to the bottom with a long tail. The mean and standard deviation of the data is 29.8s and 23s respectively. %, is far greater than for the number of waypoints they required. Indeed, there is not a strong relationship between the number of waypoints a participant required and the time they took to find the target, as showcased by participant $s7$'s wide mean waypoint count and spread in Figure~\ref{fig:nsteps-participants} and the small spread and close-to-average time to target in Figure~\ref{fig:time-participants}. Furthermore, the opposite can be observed in participant $s8$'s small spread and low mean waypoint count in Figure~\ref{fig:nsteps-participants} and high spread and high mean time to target in Figure~\ref{fig:time-participants}.

@@

In comparison to the remotely-assisted VizWiz system~\cite{bigham2010vizwiz} covered in Section~\ref{sec:previous-work} (mean 92s, standard deviation 37.7s), our results look very encouraging, although there might be significant variations in case of participants with actual vision impairments.

@@

\noindent In this work we presented and tested an MDP-based system to guide a person with vision impairments towards a target object with no a~priori knowledge of the environment. We implemented the system in an Android app that we tested with a small group of sighted people to determine its viability and effectiveness at guiding a user to manipulating a camera and finding a target object. We found that is works well when compared to alternative systems, with a respectable mean time of target acquisition of 32s. However, the system can be improved by refining the search strategy and implementing an automatic fail-state recovery system that will guide a user back to neutral position if a fail-state is entered (e.g.\ a blank section of wall). Furthermore, a different dataset, purpose-built for object spatial relations, will give the MDP a more accurate understanding of the world and an improved transition model.

The next steps to this project will involve using the data recorded here to improve and refine the MDP model that provides the guidance. The model can also be expanded to include more objects and actions. However, an important consideration and expansion to the problem is to change the MDP to an Partially Observable MDP (POMDP) model that will bring the model more in line with reality by acknowledging that the objects are not perfectly observable and humans do not perfectly follow guidance instructions. This adjustment may be very important when we replace the QR code scanner with an object recognition that are typically more prone to making classification errors. Finally, the model can be adapted to use on-line learning techniques in order to further refine its transition model as the guidance system is used.

>>>>

\keywords{Active vision, object search, visual impairment, Markov Decision Process}

\abstract{Modern smartphones can provide a multitude of services to assist people with visual impairments, and their cameras in particular can be useful for assisting with tasks, such as reading signs or searching for objects in unknown environments. Previous research has looked at ways to solve these problems by processing the camera's video feed, but very little work has been done in {\em actively} guiding the user towards specific points of interest, maximising the effectiveness of the underlying visual algorithms. In this paper, we propose a control algorithm based on a Markov Decision Process that uses a smartphone's camera to generate real-time instructions to guide a user towards a target object. The solution is part of a more general active vision application for people with visual impairments. An initial implementation of the system on a smartphone was experimentally evaluated with participants with healthy eyesight to determine the performance of the control algorithm. The results show the effectiveness of our solution and its potential application to help people with visual impairments find objects in unknown environments.}

@@

\noindent It is estimated that almost half a billion people worldwide live with mild to severe visual impairments or total blindness~\cite{bourne2017magnitude} and significant effort is being made to enable these people to lead more independent lives. Modern improvements in mobile computing power and image processing techniques have provided researchers with new and powerful tools to solve this problem. The work presented here is part of a project to assist people with visual impairments to navigate and find objects in unknown environments with the aid of a smartphone. The proposed system implements ideas from the field of active vision~\cite{Bajcsy2017}, but replaces the typical electro-mechanical actuators of a moving camera with the body (i.e.\ arm, hand) of the user holding the smartphone, as pictured in Figure~\ref{fig:system-use}, expanding upon concepts originally proposed in~\cite{bellotto2013} and~\cite{lock2017portable}.

@@

\noindent Assistive technology for people living with visual impairments is a growing research area~\cite{Manduchi2012,khoo2016multimodal}. In recent years, the increase in mobile processing power and computer vision improvements have led to research in the use of smartphone cameras to augment or enhance a user's vision and help them find objects or other points of interest. Earlier attempts at the problem involved placing special markers or barcodes around an environment, which the user then scans with a smartphone or similar mobile device~\cite{gude2013blind,iannizzotto2005badge3d,manduchi2012mobile}. This device then uses some feedback mode, e.g.\ Braille or sound, to guide the user towards the target. %The major drawbacks to this approach is the significant upfront effort required to augment entire areas with markers to a point where the system will be useful to a VI user. Furthermore, these passive systems rely on the user scanning the marker in the first place; something that can be made more complicated with markers degrading over time, changing ambient lighting, different viewing angles and marker occlusion.

Another approach is to discard tags completely and rely on computer vision to perform the object detection, something that has become more practical with recent improvements to feature detectors and deep networks~\cite{huang2017speed,redmon2016you}. SIFT and SURF-based object detectors have also been used to detect known objects, when they are in the camera's view, and to guide the user to them using sonified instructions~\cite{schauerte2012assistive}. These type of systems is more flexible than the tag-based ones, but it has the same drawback of being {\em passive}, in the sense that it relies on having the object within the camera's view in the first place. Also, no clear performance metrics are reported in the previous paper. The VizWiz system~\cite{bigham2010vizwiz} offloads the object recognition tasks to an Amazon Mechanical Turk worker who then provides feedback on where the object of interest is located relative to the user. The VizWiz has the advantage of being fairly robust and is able to classify a great deal of objects with little effort from the user and can provide natural, human-generated and curated directions. However, this approach does not enhance user independence, since a person with visual impairments is now beholden to an online worker instead of a relative, friend or bystander. Furthermore, a good internet connection is required on the device, possibly limiting its use in some poor-reception areas.

@@

In summary, much research has been conducted on recognition of and guidance towards target objects, including active vision solutions for image classifiers and robotic systems. However, to our knowledge, no previous work has been done on {\em active} object search and guidance for humans, which would especially benefit people with visual impairments. In this paper, we implement such an active vision system with a human in the loop
% ~\cite{bellotto2013,lock2017portable}
that guides the user towards an out-of-view target object. Our system exploits prior knowledge of the objects spatial distribution within an indoor environment, learned from a dataset of real-world images, and the history of past object observations made during the search.
%With this, we hope to make the searching process less random and move towards a more natural and efficient object searching process to the benefit of the VI\@.

@@

\noindent The work presented in this paper is a fundamental step towards a more general project's goal to develop a stand-alone system that can guide a person with visual impairments to his/her destination with minimal user input or intervention. A complete system diagram is given is Figure~\ref{fig:control-loop}. This closed-loop system is conceptually similar to other classical control problems, where the difference between desired and actual state of a process is used to generate a control signal that changes the process itself.

@@

In a real application for people with visual impairments, the position of the waypoint would be given to the user by a set of audio or vibrotactile instructions. However, since we are mainly interested in evaluating the control algorithm of our system and not the interface ($K$ and not $u$), our current prototype generates guidance instructions with four on-screen arrows (see Figure~\ref{fig:system-screenshot}).
% This might seem like an unnecessary extra step, but this system is designed for user with visual impairments who are limited to audio instructions, which makes it necessary to split the instructions up into separate channels that can be communicated more simply.
% This requirement becomes clear when one considers the bandwidth limitations of a person's audio channel compared to the visual one.
Obviously, this visual interface is only used for debugging and experimental evaluation of the controller, and it will be replaced by an opportune audio interface, e.g.~\cite{bellotto2013}, at a second stage.

@@

Table~\ref{tab:results-full} shows a fairly consistent spread for the TAR across the participants. The inter-participant spread in Table~\ref{tab:results-summary} ($\sigma=11\%$) is fairly significant, perhaps indicating that the user's search behaviour and strategy affects the target acquisition performance, but with an average TAR of $82\%$, it is clear that the system successfully finds the target object during the vast majority of searches.

@@

\noindent The cumulative distribution of the time it took the participants to reach the target object is given in Figure~\ref{fig:time-participants}. We see that the distribution is heavily skewed to the bottom with a long tail. The mean and standard deviation of the data is 34s and 23s respectively. %, is far greater than for the number of waypoints they required. Indeed, there is not a strong relationship between the number of waypoints a participant required and the time they took to find the target, as showcased by participant $s7$'s wide mean waypoint count and spread in Figure~\ref{fig:nsteps-participants} and the small spread and close-to-average time to target in Figure~\ref{fig:time-participants}. Furthermore, the opposite can be observed in participant $s8$'s small spread and low mean waypoint count in Figure~\ref{fig:nsteps-participants} and high spread and high mean time to target in Figure~\ref{fig:time-participants}.

@@

In comparison to the remotely-assisted VizWiz system~\cite{bigham2010vizwiz} covered in the related work (mean 92s, standard deviation 37.7s), our results look very encouraging, although there might be variations in case of participants with visual impairments.

@@

\noindent In this work we presented and tested an MDP-based system to guide a person with visual impairments towards a target object with no prior knowledge of the environment. We implemented the system in an Android app and tested it with sighted users to determine the effectiveness of the active object search algorithm. We found that it works generally well, even when compared to alternative human-guided systems. However, the solution can be improved by refining the search strategy and implementing an automatic fail-state recovery when the user points to an empty scenario, like a blank wall. Furthermore, a purpose-built dataset with clear object spatial relations would enable the creation of a more accurate transition model for the MDP controller.

Future work will include the replacement of the QR codes with a real vision-based object detector, possibly extending the number of items and actions. However, in order to consider the uncertainty introduced by such detector, the MDP will have to be replaced by a Partially Observable MDP (POMDP), taking into account that the objects are not perfectly observable and humans might not follow the guidance instructions accurately. Further directions of research include on-line learning techniques for model adaptation that better follow the user profile of each individual and the possible performance change over time.
